---
title: "Volatility Analysis of Natural Gas Prices with GJR-GARCH"
author: "Emanuelle"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: darkly
    highlight: zenburn
    toc: true
    toc_float: true
---

```{css, echo=FALSE}
pre {
  background-color: #222222 !important;
  color: #f1f1f1 !important;
  border: none;
}
```
---


```{r setup, include=FALSE}
library(httr)
library(jsonlite)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tseries)     # Para o teste ADF
library(FinTS)       # Para o teste ARCH
library(rugarch)     # Para modelos GARCH
library(tibble)

```


## 1. Data Import from EIA API

```{r}
api_key <- "F4MbCfNYB7lBrDFsbzbtGXzIo4exuULH0gQKuN0G"

series_ids <- list(
  Price            = "NG.RNGWHHD.D",
  Gross_Production = "NG.N9050US2.M",
  Storage          = "NG.NW2_EPG0_SWO_R48_BCF.W",
  Exports          = "NG.N9150US2.M",
  Imports          = "NG.N9100US2.M"
)

get_eia_series_v2 <- function(series_id, label, api_key) {
  url <- paste0("https://api.eia.gov/v2/seriesid/", series_id, "?api_key=", api_key)
  response <- GET(url)
  data <- fromJSON(content(response, "text", encoding = "UTF-8"))
  
  if (!is.null(data$response$data)) {
    df <- data$response$data
    df <- df[, c("period", "value")]
    colnames(df) <- c("Date", label)
    df$Date <- ifelse(nchar(df$Date) == 7, paste0(df$Date, "-01"), df$Date)
    df$Date <- as.Date(df$Date)
    df <- df %>% filter(Date >= as.Date("2010-01-01")) %>% arrange(Date)
    return(df)
  } else {
    warning(paste("No data for", series_id))
    return(data.frame())
  }
}

data_list <- list()
for (label in names(series_ids)) {
  df <- get_eia_series_v2(series_ids[[label]], label, api_key)
  if (nrow(df) > 0) {
    data_list[[label]] <- df
  }
}
```

---


# 2. Data Transformation: Log Returns

```{r}
data_list$Price <- data_list$Price %>%
  arrange(Date) %>%
  mutate(Log_Returns = log(Price) - log(lag(Price)))

returns <- na.omit(data_list$Price$Log_Returns)

ggplot(data_list$Price, aes(x = Date, y = Log_Returns)) +
  geom_line(color = "cyan") +
  labs(title = "Log Returns - Natural Gas Spot Price", x = "Date", y = "Log Returns") +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#222222", color = NA),
    panel.background = element_rect(fill = "#222222", color = NA),
    panel.grid.major = element_line(color = "#444444"),
    panel.grid.minor = element_line(color = "#333333"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    plot.title = element_text(color = "white", face = "bold")
  )

```

**Interpretation:** Log returns highlight clusters of volatility — a key feature in commodity markets. In practice, this means price shocks (like weather or war) ripple over time, increasing market uncertainty.

---

# 3. Stationarity Test (ADF)

```{r}
adf_result <- adf.test(returns)
adf_result
```

**Interpretation:** The p-value < 0.01 confirms the return series is stationary. This supports using GARCH-type models, which assume a stable mean and changing variance.

---

# 4. ARCH LM Test

```{r}
ArchTest(returns, lags = 12)
```

**Interpretation:** We reject the null of no ARCH effects. That means conditional heteroskedasticity is present — validating the need for models like GARCH or GJR-GARCH.

---

# 5. GJR-GARCH(1,1) with t-student Distribution

```{r}
spec <- ugarchspec(
  variance.model = list(model = "gjrGARCH", garchOrder = c(1,1)),
  mean.model = list(armaOrder = c(0,0), include.mean = TRUE),
  distribution.model = "std"
)

gjr_fit <- ugarchfit(spec = spec, data = returns)
gjr_fit
```

**Interpretation:** The t-student GJR-GARCH captures:

- High volatility persistence (beta ≈ 0.77)
- Shock responsiveness (alpha ≈ 0.22)
- Heavy tails (shape ≈ 3.5), meaning extreme events are more likely than under normality.

---

# 6. Model Diagnostics

```{r}
infocriteria(gjr_fit)

res <- residuals(gjr_fit, standardize = TRUE)
Box.test(res, lag = 20, type = "Ljung-Box")
Box.test(res^2, lag = 20, type = "Ljung-Box")
```

**Interpretation:** No autocorrelation in residuals or squared residuals → the model fits well. This makes it suitable for forecasting volatility or risk modeling in financial and energy contexts.

---

# 7. Conditional Volatility Plot

```{r}
volatility_series <- sigma(gjr_fit)
dates <- tail(data_list$Price$Date, length(volatility_series))


vol_df <- tibble(
  Date = dates,
  Volatility = volatility_series
)

ggplot(vol_df, aes(x = Date, y = Volatility)) +
  geom_line(color = "cyan", linewidth = 0.8) +
  labs(
    title = "Estimated Conditional Volatility (GJR-GARCH)",
    x = "Date", y = "Volatility"
  ) +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#222222", color = NA),
    panel.background = element_rect(fill = "#222222", color = NA),
    panel.grid.major = element_line(color = "#444444"),
    panel.grid.minor = element_line(color = "#333333"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    plot.title = element_text(color = "white", face = "bold")
  )

```

**Interpretation:** Spikes in volatility match real events (e.g. COVID-19, Ukraine War), showing the model captures real-world uncertainty well.  
With the t-student distribution, the **gamma** (asymmetry) term is not significant — meaning **fat tails** (extreme events) are more influential than the sign of the shock.

---

# 8. Final Insights and Real-World Relevance

- Natural gas price volatility is **non-constant**, with **clusters and persistence**.
- GJR-GARCH(1,1) with t-student distribution gives a **better statistical fit** than with normal distribution.
- **Asymmetry** (gamma) is significant under normality, but not under t-student — supporting the idea that **extremes matter more than direction**.
- The model supports **risk forecasting, stress testing**, and **policy planning** in energy economics.

**Applications:**
- **Hedging strategies** for LNG producers and importers.
- **Market risk modeling** by analysts and portfolio managers.
- **Policy insights** on energy security and supply volatility.

